---
title: "dsfbnine"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dsfbnine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
devtools::install_github("nineluijendijk/dsfbnine")
library(dsfbnine)
```

## Package contents

### Dataset measlesdata

The dsfbnine package contains some functions to help me with my data analysis and one cleaned dataset `measlesdata`. It contains data on measles cases and population numbers in EU countries from 2018 and 2019. The dataset loads as a dataframe with 744 rows and 8 variables. It is available via lazy-loading, meaning columns can be accessed as follows:

```{r measlesexample}
measlesdata$cases %>% head(n = 20)
```

### Function dataSummarizer()

The function `dataSummarizer()` calculates the sum, mean, median and standard deviation of data, grouped by a given argument:

```{r dataSummarizerexample}
dataSummarizer(measlesdata, groupBy = c("CountryName", "year")) %>% head(n = 10)
```

### Function illnessPlot()

The function `illnessPlot()` 

```{r illnessPlotexample}
illnessPlot(measlesdata, countries = c("Netherlands", "Belgium", "France"), years = 2020:2022, parameter = cases)
```

